\documentclass[../main.tex]{subfiles}

\begin{document}

\section{Adjoint TYMH theories}%
\label{sec:adjoint}

We start our investigation on the possible BRST and gauge symmetry breaking, recalling that the traditional Higgs action functional is $\mathrm{g}$ metric-contaminated, and not an $s$-boundary. This is not an ideal candidate to couple to TYM theories, as such coupling would explicitly break the TYM symmetries, while the naïve introduction of a scalar field would give non-vanishing contributions to the Witten index of the surviving BRST symmetry. Thus, the spontaneous symmetry breaking scenario would be completely jeopardized.

To circumvent these issues, we build upon the Higgs-like toy model first proposed by K.~Fujikawa in~\cite{fujikawa1983a}. Fujikawa's original work is not gauge invariant, so what follows is our gauge invariant generalizations of it, first in the adjoint, and later in the fundamental (and realified fundamental) representations of $SU(N)$. Albeit we focus on the coupling to TYM theories, the methodology below can be adapted to several other topological gauge field theories, regardless of their spacetime dimensions. It is also important to stress that the BRST and gauge symmetry breakings are not necessarily related to each other. Each breaking can happen independently, on its own terms --- albeit in the examples below BRST happens to follow gauge symmetry breakings.

All of that said, let us consider the field variables $\Phi (x)$, $\eta (x)$, $\xi (x)$, $B (x) \in C^{\infty} ( \mathrm{ad} P \otimes \bigwedge^0 \mathbb{R}_{\infty}^{4} )$, obeying the algebra
\begin{subequations}%
  \label{eq:afh-brst}
  \begin{align}
    s \Phi & = - \left[c, \Phi\right] + \eta \;, \\
    s \eta & = -\left[c, \eta\right]  \;,        \\
    s \xi  & = - \left[c, \xi\right] + B \;,     \\
    s B    & = -\left[c, B\right]  \;.
  \end{align}
\end{subequations}
We will refer to them as adjoint Fujikawa-Higgs (aFH) fields, and their statistics can be found in Table~\ref{tab:gradings}. Introduced as a pair of $s$-doublets, aFH fields can only build gauge invariant polynomials which are $s$-boundaries, thus living outside the $s$-cohomology groups. This is a general result in BRST algebra, known as the \textit{doublet theorem}. This theorem justifies the form of the action functional~\eqref{eq:afh-action-s-exact}, and also implies that the $s$-cohomology groups are explicitly preserved: no new observables are introduced; especially no $\mathrm{g}$ metric-contaminated ones. In other words, the topological structure of TYM theories is untouched. Moreover, $s$-doublets give vanishing contributions to the Witten index of $s$. In this way, their introduction cannot naïvely change the BRST stability or instability scenarios. % chktex 36

We naturally choose to minimally couple TYM and aFH fields, defining the Fujikawa-Higgs sector of the $SU(N)$ invariant \textquote{adjoint topological Yang-Mills-Higgs} (aTYMH) theories via the action functional
\begin{subequations}%
  \label{eq:afh-action}
  \begin{align}%
    % S_{\text{aFH}} \left[A, \psi, \Phi, \xi, \eta, B\right] & \equiv s \int \tr \left\{D \xi \star D \Phi + \xi \Phi \left[m^2 + g \left(B \Phi - \xi \eta\right)\right] \star \mathds{1} \right\} \;,                      \label{eq:afh-action-s-exact} \\
    S_{\text{aFH}} \left[ \Xi \right] & \equiv s \int \tr \left\{D \xi \star D \Phi + \xi \Phi \left[m^2 + g \left(B \Phi - \xi \eta\right)\right] \star \mathds{1} \right\} \;,                      \label{eq:afh-action-s-exact} \\
                                      & = \int \tr \left\{ -DB \star D \Phi - D \xi \star D \eta + \left( \mathrm{ad}_{\Phi}D \xi - \mathrm{ad}_{\xi} D \Phi \right) \star \psi \vphantom{m^2} \right. + \nonumber                  \\
                                      & + \left. \left[ m^2 \left( B \Phi - \xi \eta \right) + g {\left( B \Phi - \xi \eta \right)}^{2} \right] \star \mathds{1}  \right\} \;, \label{eq:afh-action-full}
  \end{align}
\end{subequations}
where $\Xi$ is shorthand for $A, \psi, \Phi, \eta, \xi, B$; $m^2$ and $g$ are mass and coupling parameters, respectively, and; $\star \mathds{1}$ is the unitary volume 4-form on spacetime\footnote{One can see that this functional is actually agnostic about the spacetime dimension.}. This is not the most general minimally coupled $s$-boundary we could have built out of aFH fields. The addition of more terms, \textit{e.g.}, $s \tr (D \xi \star DB)$, most likely benefits its renormalizability properties, but does not alter its symmetry breaking patterns. Thus, we chose not to add them. Moreover, we purposefully omitted all vertices involving $c$ from~\eqref{eq:afh-action-full}, which is equivalent to work with the already mentioned Cartan model --- isomorphic to the Kalkman model on the subspace of basic forms. It can be shown that the integrand in~\eqref{eq:afh-action-s-exact} is indeed a basic form. For convenience, we will work with the Cartan model whenever we are dealing with basic forms. Explicitly,
\begin{subequations}%
  \label{eq:cartan-model-afh-brst}
  \begin{align}
    s A    & =   \psi \;,   & s \Phi & =   \eta \;, & \hspace{200pt} \\
    s \psi & =  - D\phi \;, & s \eta & = 0 \;,      & \hspace{200pt} \\
    s \phi & = 0 \;,        & s \xi  & =  B  \;,    & \hspace{200pt} \\
    s F    & = -D\psi       & s B    & = 0      \;. & \hspace{200pt}
  \end{align}
\end{subequations}
This $s$, in particular, is nilpotent only up to the infinitesimal gauge transformations generated by $\phi$, namely, $\delta_{\phi} \equiv - s^2$. Now, we can make it clear that the integrand~\eqref{eq:afh-action-s-exact} is basic because it is independent of $\phi$, and invariant under $\delta_{\phi}$. This is the definition of a basic form, and the reason why this $s$ is nilpotent on this subspace. The aFH fields also individually fit this definition, further implying they commute with $\phi$.

We focus on the bosonic potential
\begin{equation}%
  \label{eq:potential}
  V_{\text{aFH}} \left(\Phi, B\right) = \tr \left[m^2 B \Phi + g {( B \Phi )}^2\right]
\end{equation}
with non-trivial local minima
\begin{subequations}%
  \label{eq:minima}
  \begin{align}
    B_0    & \equiv \mathrm{diag} \left(b_1, \ldots, b_N \right) \;,             \\
    \Phi_0 & \equiv \mathrm{diag} \left(\varphi_1, \ldots, \varphi_N \right) \;,
  \end{align}
\end{subequations}
where each pair of eigenvalues $(b_j, \varphi_j)$ must solve
\begin{subequations}%
  \label{eq:eigenvalues-conditions}
  \begin{align}
    m^2 b_j + 2g b_j^2 \varphi_j +\alpha        & = 0 \;,  \\
    m^2 \varphi_j + 2g b_j \varphi_j^2  + \beta & = 0 \;,  \\
    \textstyle\sum_{j} b_j                      & = 0 \;,  \\
    \textstyle \sum_j \varphi_j                 & = 0 \;,  \\
    \Delta_k                                    & > 0  \;.
  \end{align}
\end{subequations}
Einstein sum convention is not assumed in~\eqref{eq:eigenvalues-conditions}, and indexes run as $j \in \left\{ 1, \ldots, N \right\} $, $k \in \left\{ 5, \ldots, 2+2N \right\} $. The parameters $\alpha, \beta \in \mathbb{R}$ are Lagrange multipliers enforcing the traceless condition characteristic of $\mathfrak{su}(N)$ matrices, and $\Delta_k$ is the determinant of the leading $k \times k $ principal minor of the bordered Hessian of $V_{\text{aFH}}$. We stress these are sufficient, but not strictly necessary conditions for local minima. For instance, if $N=2$, then only one independent Casimir operator exists, implying $\tr [ {(B \Phi)}^2 ] \propto { [ \tr ( B \Phi ) ] }^2$. In such case, solving~\eqref{eq:eigenvalues-conditions} directly can be completely avoided by simply \textquote{completing the square} in~\eqref{eq:potential}.

We find that the gauge symmetry breaking pattern is exactly the same as the one in the adjoint Higgs from particle physics. $(B_0, \Phi_0)$ are always valued in the Cartan subalgebra $\bigoplus^{N-1} \mathfrak{u}(1)$ of $\mathfrak{su}(N)$, and have stabilizer $\mathfrak{h}_0 = \left\{ X \in \mathfrak{su}(N) \; ; \; \mathrm{ad}_{B_0} X = 0, \;\mathrm{ad}_{\Phi_0} X = 0\right\} $ in the minima moduli $\mathcal{V}_0$. In the trivial neighborhood, $\mathfrak{h}_0 \sim \mathfrak{su}(N)$. Away from it, gauge symmetry is spontaneously broken, $\mathfrak{h}_0 \subset \mathfrak{su}(N)$. Specifically, in a region where the multiplicity of $(b_j, \varphi_j)$ is $n_j=1 \; \forall \; j$ (non-degeneracy), then $\mathfrak{h}_0 \sim \bigoplus^{N-1} \mathfrak{u}(1)$, implying $N(N - 1)$ broken generators, thus $N(N-1)$ massive gauge field components. In a region where $l \in \left\{1, \ldots, m\right\} $ labels distinct pairs of eigenvalues, and $1 < n_l \leq N \; ; \; \sum_l n_l = N$ (some degeneracy), then $\mathfrak{h}_0 \sim \bigoplus_l \mathfrak{su}(n_l) \oplus \bigoplus^{m-1} \mathfrak{u}(1)$, implying $N^2 - \sum_l n_l^2$ broken generators, thus $N^2 - \sum_l n_l^2$ massive gauge fields components. The region of complete degeneracy ($m=1$) is the trivial neighborhood, implying no broken generators, no massive gauge field components, no gauge symmetry breaking.

The action functional~\eqref{eq:afh-action} assumes the following effective form in the neighborhood of $(B_0, \Phi_0)$,
\begin{subequations}%
  \label{eq:afh-eff-action}
  \begin{align}%
    \bar{S}_{\text{aFH}} \left[ \bar{\Xi} \right] & = \bar{s} \int \tr \left\{ D \xi \star D\varphi + \mathrm{ad}_{\Phi_0} D \xi \star A + \xi ( \Phi_0 + \varphi ) \left[ m^2 + g ( B_0 \Phi_0 \right. + \right.                                                      \nonumber   \\
                                                  & \left. + \left.  B_0 \varphi + b \Phi_0 + b \varphi - \xi \eta ) \vphantom{m^2} \right] \star \mathds{1} \right\}     \;,                \label{eq:afh-eff-action-exact}                                                       \\
                                                  & = \int \tr \left\{ \vphantom{m^2} \mathrm{ad}_{B_0} \mathrm{ad}_{\Phi_0} A \star A + \left( \mathrm{ad}_{B_0} \varphi + \mathrm{ad}_{\Phi_0} b \right) D \star A  + \mathrm{ad}_{\Phi_0} \xi D \star \psi \right. +  \nonumber \\
                                                  & - Db \star D \varphi - D \xi \star D \eta - \left( \mathrm{ad}_{B_0} \mathrm{ad}_{A} \varphi + \mathrm{ad}_{\Phi_0} \mathrm{ad}_{A} b \right) \star A + ( \mathrm{ad}_{\varphi} D \xi \; +  \nonumber                          \\
                                                  & - \mathrm{ad}_{\xi} D \varphi + 2 \mathrm{ad}_{\Phi_0}\mathrm{ad}_{A} \xi ) \star \psi + \left[ ( m^2 + 2g B_0 \Phi_0 ) ( b \varphi - \xi \eta ) + g \left( B_0 \varphi \right. + \right. \nonumber                            \\
                                                  & \left. + \left. \left.  b \Phi_0 + b \varphi - \xi \eta \right)^2 \right] \star \mathds{1} \right\} \;, \label{eq:afh-eff-action-full}                                                                                         % chktex 3
  \end{align}
\end{subequations}
where $\bar{S}_{\text{aFH}} [ \bar{\Xi} ] \equiv S_{\text{aFH}} [A, \psi, \Phi_0 + \varphi, \eta, \xi, B_0 + b]$; $\bar{\Xi}$ is shorthand for $A, \psi, \varphi, \eta, \xi, b$, and; $\bar{s}$ is given by~\eqref{eq:afh-bar-brst}. In particular, its quadratic part
\begin{align}%
  \label{eq:afh-eff-quad-action}
  \bar{S}_{\text{aFH}}^{(2)} \left[ \bar{\Xi} \right] & = \int \tr \left\{ \vphantom{m^2} \mathrm{ad}_{B_0} \mathrm{ad}_{\Phi_0} A \star A + \left( \mathrm{ad}_{B_0} \varphi + \mathrm{ad}_{\Phi_0} b \right) d \star A + \mathrm{ad}_{\Phi_0} \xi d \star \psi \right. +  \nonumber \\
                                                      & - \left. db \star d\varphi - d\xi \star d\eta +\left[ ( m^2 + 2g B_0 \Phi_0 ) \left( b \varphi - \xi \eta \right) + g {\left( B_0 \varphi + b \Phi_0 \right)}^2 \right] \star \mathds{1} \right\}
\end{align}
makes very clear that some of the effective gauge field components have mass defined by the mass matrix $\mathrm{ad}_{B_0}\mathrm{ad}_{\Phi_0}$. This is, of course, a direct reflection of the spontaneous gauge symmetry breaking just described. The terms proportional to $d \star A$ and $d \star \psi$ are removable via gauge fixing. And, the massive bosonic pair $(b, \varphi)$ plays the role similar to the Higgs boson field, with $(\xi,\eta)$ being its massive fermionic partner. Notice that $(\xi,\eta)$ kept its mass because the factor $m^2+2gB_0\Phi_0$ did not identically vanish thanks to the \textquote{adjointness} of our fields --- something what will change in the (realified) fundamental case.

One can easily check that no massless fermions are present in~\eqref{eq:afh-eff-quad-action}, which indicates that the topological BRST symmetry is unbroken at tree-level. The effective BRST
\begin{subequations}%
  \label{eq:afh-bar-brst}
  \begin{align}
    \bar{s} A    & =   \psi \;,   & \bar{s} \varphi & =   \eta \;,    & \hspace{200pt}                            \\
    \bar{s} \psi & =  - D\phi \;, & \bar{s} \eta    & = 0 \;,         & \hspace{200pt}                            \\
    \bar{s} \phi & = 0 \;,        & \bar{s} \xi     & =  B_0 + b  \;, & \hspace{200pt}\label{eq:afh-bar-brst-ssb} \\
    \bar{s} F    & = -D\psi       & \bar{s} b       & = 0      \;,    & \hspace{200pt}
  \end{align}
\end{subequations}
indeed remains nilpotent (up to $\delta_{\phi}$), and isomorphic to the original~\eqref{eq:cartan-model-afh-brst}. And, the effective Higgs-phase action~\eqref{eq:afh-eff-action-full} can indeed be written as the $\bar{s}$-boundary~\eqref{eq:afh-eff-action-exact}, automatically making it a $\bar{s}$-cycle\footnote{The use of the Cartan model is justified because the integrand in~\eqref{eq:afh-eff-action-exact} is basic. This is somewhat surprising given to the explicit presence of $A$ outside $D$. We attribute this to the tree-level stability of the BRST\@. As long as $s(B_0, \Phi_0) = (0,0)$, the basicness of $(B,\Phi)$ gets passed down to $(B_0, \Phi_0)$, and to $(b,\varphi)$, ultimately ensuring the basicness of the integrand (and the nilpotency of $\bar{s}$ on it).}. Clearly, the original topological features are still present in the classical realm. Nevertheless, if $ \langle 0 \rvert b \lvert 0 \rangle = 0$, then~\eqref{eq:afh-bar-brst-ssb} implies $\langle 0 \rvert \bar{s} \xi \lvert 0 \rangle = B_0 $. As long as $B_0 \neq 0$, this is a sufficient condition for spontaneous BRST symmetry breaking. In turn, Nambu-Goldstone fermionic states must necessarily be present~\cite{fujikawa1983a,witten1982a,witten1981a,salam1974a}. Since such degrees of freedom are absent at tree-level, they must be generated by loop corrections or non-perturbative effects. Saying it differently, the lack of a Nambu-Goldstone fermion in~\eqref{eq:afh-eff-quad-action} implies that the Higgs-phase aTYMH theories must be quantum mechanically non-trivial. This is a remarkably different scenario from what we started and, interestingly, one which blurs the line between spontaneous and dynamical symmetry breakings. Ultimately, we can conclude that the $\mathcal{G}$-equivariant cohomology of $\mathcal{A}$ no longer fully captures the complete set of Higgs-phase physical states: we are no longer dealing with cohomological TQFTs.

It is also worth analyzing the energy functional evaluated from~\eqref{eq:afh-action},
\begin{subequations}%
  \label{eq:afh-energy}
  \begin{align}
    \mathbb{E}_{\text{aFH}} \left[ \Xi ; r\right] & =  \int_{B^3_r} \tr \left\{ \mathcal{L}_{\lambda}c \star ( \mathrm{ad}_{\Phi} D \xi - \mathrm{ad}_{\xi} D \Phi) + s \left[ \vphantom{m^2} \mathcal{L}_{\lambda} \Phi \star D \xi + D \xi (\lambda\rfloor \star D \Phi) \right. + \right. \nonumber \\
                                                  & \left. + \left. \mathrm{ad}_\xi \star D \Phi (\lambda \rfloor A) + \xi \Phi [ m^{2} + g ( B \Phi - \xi \eta ) ] ( \lambda \rfloor \star \mathds{1} ) \right] \right\}                                                                              \\
                                                  & =  \int_{B^3_r}\tr \left[ \vphantom{m^2} \mathcal{L}_{\lambda} B \star D \Phi - \mathcal{L}_{\lambda} \Phi \star DB + \mathcal{L}_{\lambda} \eta \star D \xi - \mathcal{L}_{\lambda} \xi \star D \eta \right. + \nonumber                          \\
                                                  & + \left. ( \mathrm{ad}_{\Phi} \mathcal{L}_{\lambda} \xi + \mathrm{ad}_{\xi} \mathcal{L}_{\lambda} \Phi ) \star \psi - (\lambda \rfloor \mathfrak{L}_{\text{aFH}}) \vphantom{m^2} \right] \;,
  \end{align}
\end{subequations}
where $\mathcal{L}_{\lambda} \equiv d \lambda \rfloor + \lambda \rfloor d$ is the Lie derivative along $\lambda$, and; $\mathfrak{L}_{\text{aFH}}$ is the Lagrangian 4-form in~\eqref{eq:afh-action-full}. This is clearly a bulk trivial quantity since the non-exact piece --- the one containing two $c$ vertices --- is removable via gauge fixing\footnote{The canonical energy-momentum 3-form of a non-Abelian gauge theory generally does not originate from a basic form. This makes the Kalkman model unavoidable, resulting in the resurrection of $c$. However, if temporal gauge conditions ($\lambda \rfloor A = \lambda \rfloor \psi =0$) are assumed, then $\mathcal{L}_{\lambda}c$ identically vanishes in the bulk.}. In this sense, the situation around the trivial vacuum is quite similar to pure TYM theories, as every classical field configuration in the bulk is part of the vacua moduli. Unsurprisingly, the same can be said about the effective energy functional $\bar{\mathbb{E}}_{\text{aFH}}$, evaluated from~\eqref{eq:afh-eff-action}. Nonetheless, due to the BRST instability at quantum level, we fully expect $\langle 0 \rvert \bar{\mathbb{E}}_{\text{aFH}} \lvert 0 \rangle $ to be bulk non-trivial. From a perturbative QFT perspective, this scenario very closely resembles the Coleman-Weinberg mechanism~\cite{coleman1973a}. And, exemplifies the non-topological effective nature of the Higgs-phase quantum aTYMH theories.

Finally, we close the adjoint case making a few comments about the presence of solitonic degrees of freedom around $(B_0, \Phi_0)$. Recalling the adjoint Yang-Mills-Higgs model from particle physics, the inequivalent gauge vacua form the moduli of flat connections --- homotopically equivalent to a point. Thus, the homotopy of the adjoint Higgs sector alone, here $\pi_2 (SU(N)/{U(1)}^{m-1}) \sim \mathbb{Z}^{m-1}$, is sufficient to determine that $m-1$ independent topological solitons of spatial co-dimension 3 stabilizes in $\mathbb{R}^4_{\infty}$. A Riemannian or Lorentzian structure then implies these are vertices or monopoles, respectively. However, at the tree-level of aTYMH theories, the effective gauge vacua form the entire gauge moduli $\mathcal{A}_\mathcal{H}/\mathcal{H}$ --- most definitely not homotopic to a point. Contributions from all these vacua can potentially destabilize Higgs sector solitons, and \textit{vice versa}. Thus, a careful analysis of the complete moduli $[\mathcal{A}_{\mathcal{H}} \times C^{\infty} (\mathrm{ad}P_{\mathcal{H}})] / \mathcal{H} $ seems to be unavoidable in the classical realm\footnote{This is a very hard problem, beyond the scope of this work.}. Notwithstanding, the quantum realm does suffer from BRST instability, implying $\langle 0 \rvert \bar{\mathbb{E}}_{\text{aFH}} \lvert 0 \rangle \neq 0 $ in the bulk. This latter fact localizes the gauge vacua back onto the moduli of flat connections, ensuring that the Higgs sector vertices/monopoles are quantum mechanically stable, and present in the physical spectrum of these effective theories.

\begin{table}[htpb]
  \caption{Gradings of all TYM fields, and the representation-independent FH fields.}%
  \label{tab:gradings} \begin{tabular}{cccccccccccc} \toprule
    Field      & $A$ & $F$  & $c$ & $\psi$ & $\phi$ & $\Phi$ & $\eta$ & $\xi$ & $B$  \\
    \midrule
    Form rank  & 1   & 2    & 0   & 1      & 0      & 0      & 0      & 0     & 0    \\
    Ghost no.  & 0   & 0    & 1   & 1      & 2      & 0      & 1      & -1    & 0    \\
    Statistics & odd & even & odd & even   & even   & even   & odd    & odd   & even \\
    \bottomrule
  \end{tabular}
\end{table}

\end{document}
