\documentclass[../main.tex]{subfiles}
\begin{document}

\section{Introduction}%
\label{sec:introduction}

Dynamical BRST\footnote{Becchi-Rouet-Stora-Tyutin.} symmetry breaking is an exceptionally rare phenomenon in quantum field theory. This is largely because the nilpotency of the BRST charge $Q$, and its overall cohomological structure, is protected by a topological invariant of the field space --- the Witten index $\tr {(-1)}^F$~\cite{witten1981a,witten1982a,fujikawa1983a}. A non-vanishing index guarantees the balance between bosonic and fermionic zero modes, ensuring that the BRST symmetry remains unbroken. Physical observables can then be consistently defined in the classical and quantum realm through its cohomology groups. Not coincidently, most consistent BRST invariant quantum field theories are defined in the $Q \lvert 0 \rangle = 0$ regime.

Historically, any sign of BRST symmetry breaking was regarded as a pathology --- an indication that the gauge fixing and/or quantization procedure had failed to properly isolate the subspace of physical states. However, the modern understanding has reshaped this perspective. In the Gribov–Zwanziger framework, for instance, the BRST instability --- made explicit by the restriction of gauge field configurations to the Gribov region ---, is interpreted as a possible hint of the color confinement in non-Abelian gauge theories~\cite{gribov1978a,zwanziger1989a,maggiore1994a,sorella2009a,dudal2012a,vandersickel2012a}\footnote{Recent developments have shown that the breaking is not catastrophic, and Gribov-Zwanzinger theory can be recast in a new BRST invariant form in several different choices of gauge functions~\cite{pereira2015a,capri2015a,pereira2016a,pereira2016b,pereira2016c}.}. Similarly, in the Parisi–Sourlas framework of stochastic dynamics, unstable BRST symmetries can be related to the emergence of chaos, long-time correlations, $1/f$ noise, \textit{etc.}~\cite{parisi1982a,ovchinnikov2016a,ovchinnikov2025a}. What was once considered a breakdown in consistency, is now understood as a window into non-perturbative phenomena.

In this work, we analyze spontaneous BRST and gauge symmetry breaking in a class of topological quantum field theories (TQFTs) known as \textquote{topological Yang-Mills} (TYM)\footnote{There is no standard use of the term \textquote{TYM} in the literature. It can mean field theories whose Lagrangian density is $\tr (F \wedge F)$, also referred to as Baulieu-Singer theories~\cite{baulieu1988a}. But, it can also mean twisted $\mathcal{N}=2$ Super-Yang-Mills theories, also referred to as Donaldson-Witten theories~\cite{witten1988d,witten1991a}. We adopt the former meaning (a comparison between the two can be found in~\cite{junqueira2021a}).}. As we briefly review in Section~\ref{sec:tym}, these are $SU(N)$ gauge invariant topological field theories, whose BRST quantization give exact path integral representations for the so-called Donaldson polynomials --- smooth invariants of 4-manifolds, carrying global information about differentiable structures~\cite{donaldson1983a,donaldson1990a,donaldson1990b,witten1988d,baulieu1988a,witten1991a}. Since the most current experimental and observational data support the existence of four large physical dimensions~\cite{lee2020a,wenhai2020a,chakravarti2020a,visinelli2018a,vagnozzi2019a}, TYM theories acquire immediate relevance in our understanding of gravity, and the large-scale structure of spacetime~\cite{witten1988a,brans1992,maluga1996,maluga2007}.

TYM theories, remarkably, also have relevance in the non-perturbative sector of quantum chromodynamics (QCD). This is because Donaldson polynomials are intrinsically related to intersection forms on the Yang-Mills (YM) instanton moduli. These are precisely the non-perturbative field configurations responsible for the huge degeneracy of the QCD vacuum. In fact, the instantonic vacuum-to-vacuum amplitudes can change the chirality of fermions, explaining the $U(1)$ axial anomaly and, consequentially, the huge mass of the $\eta '$ meson~\cite{thooft1976a,thooft1986a}. Clearly, it is not hard to image that results in TYM theories can, potentially, trickle down to phenomenological implications in particle physics.

TYM theories are cohomological field theories (Witten-type TQFTs). This means their defining symmetry is a \textquote{topological BRST symmetry}: the BRST operator is the differential for an equivariant cohomology model of a field space. On one hand, this is what enforces the subspaces of physical states to be exclusively populated by BRST invariant vacua ($\exists$ no local excitations), also constraining the set of observables to be purely topological in nature. On the other, it obviously hinders any naïve attempts to directly connect TYM theories with traditional local physics.

Interestingly enough, TYM theories have vanishing Witten index. Its topological BRST symmetry is unprotected, and possibly unstable. This is precisely the focus of our investigation. Long story short, we have found that, when adequately coupled to \textquote{Higgs-like fields}, these theories exhibit BRST instability with the emergence of mass scales, revealing non-topological Higgs-like phases. In particular, with all the fields in the adjoint representation of $SU(N)$, BRST is unstable beyond the tree-level --- implying that the effective theories have non-trivial quantum mechanical regimes generically populated with local degrees of freedom: massless and massive gauge field states, and Nambu-Goldstone fermionic field states. There is also the widespread presence of non-local field states associated with stable topological solitons. With the Higgs-like fields  in the fundamental representation, we have found that no spontaneous symmetry breaking occurs unless we appeal to realification. The realified $SU(N) \hookrightarrow SO(2N)$ theories, strikingly, have instability and emergence of a physical mass scale at tree-level. Their classical Higgs-phase dynamics exhibit, explicitly, the presence of the Nambu-Goldstone fermion, while stable topological solitons are generically absent --- just a single vertex/monopole is present in $N=2$. Anyhow, the adjoint case might be an interesting realization of the Coleman-Weinberg mechanism~\cite{coleman1973a}, while the realified fundamental case might inspire more direct connections between high energy topological gravity models and low energy geometrodynamics~\cite{sako1997a,mielke2011a,alexander2016a,sadovski2017a,chengzheng2017a,agrawal2020a,edery2023a,tianyao2023a,kehagias2021a,raitio2024a,sadovski2025a}.

This work is structured as follows. In Section~\ref{sec:tym}, we define our conventions, establish the mathematical framework, and review the main introductory features of TYM theories. The mathematical framework is quite rigorous, but the universal bundle construction is one of the quickest ways to introduce several aspects of these theories. In Section~\ref{sec:adjoint}, we develop the so-called \textquote{Fujikawa-Higgs sector}, with all fields in the adjoint representation of $SU(N)$. We evaluate the Higgs-phase effective theories, argue for the BRST instability beyond tree-level, and describe the effective field states. In Section~\ref{sec:fundamental}, we follow very closely the methodology employed in Section~\ref{sec:adjoint}, but with the \textquote{Fujikawa-Higgs fields} in the fundamental representation. Realification follows, we evaluate the realified Higgs-phase effective dynamics, argue for the tree-level BRST instability, and describe the effective degrees of freedom present. Finally, in Section~\ref{sec:conclusions}, we present our concluding remarks.

\end{document}
