\documentclass[../main.tex]{subfiles}

\begin{document}

\section{Topological Yang-Mills theories}\label{sec:tym}

Let spacetime be $\mathbb{R}^{4}_{\infty}$, endowed with a globally flat metric $\mathrm{g}$. The 4-manifold $\mathbb{R}^{4}_{\infty}$ is the one-point compactification of the standard $\mathbb{R}^4$, compatible with its conformal structure at infinity. This is a convenient choice of background for TYM, because it is compact, Riemannian and 4-dimensional --- necessary conditions for the rigorous definition of the YM instanton\footnote{$\mathbb{R}^{4}_{\infty}$ is also homeomorphic, but not diffeomorphic to $S^4$. Additionally, $\mathrm{g}$ is conformally equivalent to a round metric on the 4-sphere. Thus, CFTs on $\left( \mathbb{R}^{4}_{\infty}, \mathrm{g}\right)$ can pick up non-trivial topological effects while conveniently defined on a flat background.}. Due to the nature of topological field theories, we prefer to work with differential forms. For future reference, $\wedge$ is the exterior product, $d$ is the exterior derivative, $v \rfloor$ is the interior product along the smooth vector field $v$, and $\star$ is the Hodge (star) dual, defined with respect to $\mathrm{g}$. All these operations occur on the exterior algebra $\bigwedge \mathbb{R}^{4}_{\infty}$ of spacetime.

Let $SU(N)$, with Lie algebra $\mathfrak{su}(N)$, be our structure group. This is a physically relevant choice, but also mathematically convenient as $SU(N)$ YM instantons are reasonably well-understood\footnote{The moduli of irreducible $SU(N)$ YM instantons is finite-dimensional, generically smooth and Kähler. Additionally, $SU(N)$ YM instantons can be fully classified via the ADHM construction. The same cannot be said about other choices of symmetry group.}. Also, for future reference, we define on each event $x \in \mathbb{R}^{4}_{\infty}$ the $\mathbb{Z}_2$ graded Lie bracket $[X,Y] \equiv X \wedge Y-{(-1)}^{|X||Y|}Y \wedge X$, where $|X|$ is the grading (statistics) of $X$; the adjoint representation $\mathrm{ad}_X Y \equiv \left[X, Y\right]$, and; the Killing form $\kappa \left(X,Y\right) \equiv \tr \left( \mathrm{ad}_{X} \mathrm{ad}_{Y} \right)$ acting on $X, Y \in \mathfrak{su}(N) \otimes \bigwedge_x \mathbb{R}^{4}_{\infty}$.

The global framework of an $SU(N)$ gauge theory is that of the universal bundle $[ SU(N)\times\mathcal{G} ] \hookrightarrow ( P \times \mathcal{A} ) \rightarrow ( \mathbb{R}^{4}_{\infty} \times \mathcal{M} )$~\cite{baulieu1985a,baulieu1988a}. The smooth  $(N^2+3)$-manifold $P$ is part of the more commonly known principle bundle structure $SU(N) \hookrightarrow P \rightarrow \mathbb{R}^{4}_{\infty}$, together with its adjoint bundles, $ \mathrm{Ad} P \equiv P \times_{SU(N)} SU(N) $ and $ \mathrm{ad}P \equiv P \times_{SU(N)} \mathfrak{su}(N)$. The space of $SU(N)$ connections on $ P $ is $ \mathcal{A} \equiv C^{ \infty } ( J^1 P ) $, where $ J^1 P $ is the 1st jet bundle of $P$, and; the space of smooth gauge transformations is $ \mathcal{G} \equiv C^{ \infty } ( \mathrm{Ad}P ) $, where $ \mathrm{Lie} (\mathcal{G}) \equiv C^{ \infty } ( \mathrm{ad} P ) $ is its Lie algebra. Finally, $ \mathcal{M} \equiv \mathcal{A}/\mathcal{G}$ is the gauge moduli\footnote{The definition of a universal bundle requires a free action of $SU(N) \times \mathcal{G}$ on $P \times \mathcal{A}$. Thus, it is implicit that we are only considering framed irreducible connections, preventing singularities in $\mathcal{M}$ and the ill-posedness of its cohomological problem.}.

A gauge field $A(x)$ is defined as an element in $C^\infty ( \mathrm{ad}P \otimes \bigwedge^1\mathbb{R}^{4}_{\infty} )$, which results from a pullback to spacetime of an $SU(N)$ connection living on $P$. Similarly, a universal gauge field $\tilde{A}(x)$ is defined as the result of a pullback to $\mathbb{R}^{4}_{\infty} \times \mathcal{M}$ of a universal $[SU(N) \times \mathcal{G}]$ connection living on $P \times \mathcal{A}$. They are both related by
\begin{equation}
  \label{eq:universal-connection}
  \tilde{A} = A + c \;,
\end{equation}
where $c(x) \in C^\infty ( \mathrm{ad}P \otimes \bigwedge^0\mathbb{R}^{4}_{\infty} )$ is a projection of the Maurer-Cartan form on $ \mathrm{ Lie } (\mathcal{G})$ --- the well-known Faddeev-Popov ghost field. The universal exterior derivative $\tilde{d}$ can also be expressed accordingly,
\begin{equation}%
  \label{eq:universal-exterior-derivative}
  \tilde{d} = d + s \;,
\end{equation}
where $s$ is a projection of the exterior derivative on $ \mathrm{Lie} (\mathcal{G})$ --- the well-known BRST operator.

The $\mathbb{Z}_2$ grading we mentioned earlier is actually a $\mathbb{Z}_2$ bi-grading that exists in the exterior algebra on $P \times \mathcal{A}$, but which descends to spacetime as the differential form rank, and ghost number. Thus, equations~\eqref{eq:universal-connection} and~\eqref{eq:universal-exterior-derivative} is just a bi-grading split into \mkbibquote{components} $(1,0)$ and $(0,1)$. In particular, they give the meaning of an 1-form with ghost number 0 to $A$, and a 0-form with ghost number 1 to $c$. Accordingly, the statistics of $A$, and $c$, is odd (fermionic) --- check Table~\ref{tab:gradings} for reference. By definition, $\tilde{d}$, $d$, and $s$, are nilpotent operators, implying fermionic nature, and resuming~\eqref{eq:universal-exterior-derivative} to $sd + ds = 0$.

The universal curvature $ \tilde{F} $ of $\tilde{A}$ is given by\footnote{From now on, whenever the context is sufficiently clear, we will omit $\wedge$. In particular, $X \wedge X = XX = X^2$.}
\begin{subequations}%
  \label{eq:universal-curvature}
  \begin{align}
    \tilde{F} & \equiv \tilde{d}\tilde{A}+\tilde{A}^2 \;, \\
              & = F+\psi+\phi \;,
  \end{align}
\end{subequations}
and it has curvature $F\equiv dA+A^2$ of $A$ as the $(2,0)$  component; the so-called \emph{topological ghost field} $\psi \equiv sA + Dc$ as the $(1,1)$ component, and; the \emph{2nd generation topological ghost field} $\phi \equiv sc + c^2$ as $(0,2)$ component. We must clarify $D \equiv d+ \left[A, \hphantom{A}\right] $ is the exterior covariant derivative in the adjoint representation.

The Bianchi identities for $\tilde{F}$, and $F$, give
\begin{subequations}%
  \label{eq:universal-bianchi-identity}
  \begin{align}
    \tilde{D}\tilde{F}          & =0 \;,  \\
    sF+D\psi+ \left[c, F\right] & = 0 \;.
  \end{align}
\end{subequations}
And, from them, we can read the so-called \textit{TYM BRST symmetries}~\cite{baulieu1988a},
\begin{subequations}\label{eq:tym-brst}
  \begin{align}
    sA    & = -Dc + \psi \;,                     \label{eq:tym-brst-A}    \\
    sc    & = - c^2 + \phi \;,                   \label{eq:tym-brst-c}    \\
    s\psi & = -D\phi -  \left[c, \psi\right]  \;,\label{eq:tym-brst-psi}  \\
    s\phi & = - \left[ c, \phi \right]  \;,       \label{eq:tym-brst-phi} \\
    sF    & = -D\psi - \left[ c, F \right]  \label{eq:tym-brst-F} \;.
  \end{align}
\end{subequations}
This is the most general realization of the BRST algebra, capturing the full cohomological structure of traditional gauge theories\footnote{It is possible to augment this algebra, and the universal bundle construction, by considering the anti-BRST transformations~\cite{birmingham1991b,carvalho1992a,perry1993a}. However, no extra cohomological information is gained.}. In mathematical language,~\eqref{eq:tym-brst} is called the \emph{BRST/Kalkman model} of the $\mathcal{G}$-equivariant cohomology of $\mathcal{A}$. Imposing $c=0$ in~\eqref{eq:tym-brst-A}, \eqref{eq:tym-brst-psi}, and~\eqref{eq:tym-brst-F}, we get the so-called \emph{Weil model}, and imposing $c=0$ everywhere, we get the \emph{Cartan model} --- Witten's original \textquote{BRST-like supersymmetry}~\cite{witten1988d,ouvry1989a,kanno1989a,kalkman1993a}. Most importantly, these models are isomorphic to each other in the subspace of basic forms\footnote{The definition of a basic form will be more conveniently given in the next Section.}~\cite{baal1990a,witten1991a,cordes1995a}. % chktex 2

Physically,~\eqref{eq:tym-brst} represents a much stronger set of symmetries than the usual YM BRST from particle physics --- the latter can be obtained from the former by imposing the so-called \emph{horizontal conditions}, $\psi=\phi=0$. Additional two very important aspects of~\eqref{eq:tym-brst} are: (i) as already mentioned in the introduction, its Witten index vanishes --- it is topologically unprotected, and; (ii) also already mentioned, the cohomology groups it defines forbids the presence of $ \mathrm{g} $ metric-contaminated observables, including the traditional YM Lagrangian density $ \tr \left( F \star F \right) $\footnote{Implicitly $\mathrm{g}$ metric-contaminated due to the explicit presence of $\star$.}. Assuming BRST stability, all non-trivial local observables are elements in the $s$-cohomology modulo $d$-boundaries. Therefore, the only ones allowed are
\begin{equation}
  \label{eq:tym-observables}
  \mathcal{O}_n = \tr \left( \tilde{F}^n \right) \; ; \; n \in {\mathbb{N}}^{\ge 1} \;.
\end{equation}
These are invariant polynomials of $\tilde{F}$ which give Chern classes of the universal bundle. In other words, \textquote{topological BRST symmetries} force all local observables to be topological in nature. In particular,
\begin{equation}
  \label{eq:donaldson-polynomials}
  \mathcal{O}_2 = \tr \left[ F^2 + 2\psi F + \left( 2\phi F + \psi^2 \right) + 2\psi \phi + \phi^2\right]
\end{equation}
contains precisely the Donaldson polynomials first evaluated in the seminal works of S.~K.~Donaldson~\cite{donaldson1983a,donaldson1990a}, and E.~Witten~\cite{witten1988d} --- albeit for the case $N=2$\footnote{$N>2$ represent generalizations of the original polynomials, reflecting the slightly more involved nature of $SU(N>2)$ bundles over $\mathbb{R}^4_{\infty}$.}.

Among all allowed observables in~\eqref{eq:tym-observables}, only the (4,0) component of~\eqref{eq:donaldson-polynomials} is suitable for a Lagrangian density in dimension four. Thus, the TYM symmetries force the action functional to be
\begin{equation}%
  \label{eq:tym-action}
  S_{\text{TYM}}\left[ A \right]  \equiv  \int \tr\left(F^2\right)
  = 8 \pi^2 k \;,
\end{equation}
where $ k \in \mathbb{Z} $ is the \textit{instanton number}\footnote{The instanton number is generally non-vanishing on $\mathbb{R}^{4}_{\infty}$. It is well-known that $F^2$ is a global $d$-cycle. Poincaré lemma then implies $F^2$ is a local $d$-boundary. In fact, $F^2 = d \omega^{(3)} $, where $\omega^{(3)} \equiv AdA+\tfrac{2}{3}A^3$ is the famous Chern-Simons 3-form. Nonetheless, the 4th de Rham cohomology group of $\mathbb{R}^{4}_{\infty}$, namely $ H^4_{\text{dR}} \left( \mathbb{R}^{4}_{\infty}\right) \sim H^4_{\text{dR}} \left( S^{4}\right) \sim \mathbb{Z} $, obstructs $F^2$ to be a global $d$-boundary, \textit{i.e.}, under an integral sign $F^2$ does not equate $d \omega^{(3)}$:~\eqref{eq:tym-action} does not generically vanish.} of $A$. More rigorously, $k$ is the 2nd Chern number of the $SU(N)$ bundle over $\mathbb{R}^{4}_{\infty}$, on which the principal connection associated with $A$ was originally defined. This is an important clarification because $k$ classifies these bundles. Thus, gauge fields (modulo gauge transformations) are in one-to-one correspondence with them (modulo bundle isomorphisms). In this sense, TYM theories are explicitly topological, and~\eqref{eq:tym-observables},~\eqref{eq:donaldson-polynomials}, and~\eqref{eq:tym-action} are, explicitly, smooth bundle invariants.

Unsurprisingly, the field equations are trivial ($0=0$), signaling lack of local bulk dynamics. Many topological field theories can be formulated as fully extended functorial field theories~\cite{atiyah1988a,segal1988a,baez1995a,schreiber2009a,baez2009a}. In this context, their non-local bulk (or boundary) dynamics can be roughly understood as the propagation and scattering of smoothly embedded fully extended cobordisms. Moreover, TYM theories have vanishing (pure gauge) energy functionals on the bulk, but not necessary on the celestial sphere at infinity
\begin{equation}%
  \label{eq:tym-energy}
  \lim_{ r \to \infty}  \mathbb{E}_{\text{TYM}} [A; r] = \int_{S^{2}_{\infty}} \tr \left[2\left(\lambda \rfloor A\right)F\right] \;,
\end{equation}
where $\lambda$ is the globally constant vector field generating spacetime translations, and $S^{2}_r \equiv \partial B^3_r$ is the 2-sphere boundary of the 3-ball $B^3_r \subset \mathbb{R}^{4}_{\infty} $ of radius $r$. In this sense, every bulk gauge field configuration amounts to the gauge vacua.

Partition functions for TYM theories, formally defined in the weakly coupled regime, require~\eqref{eq:tym-action} to be gauge fixed. Quantum TYM (QTYM) theories have been studied in several different gauge choices~\cite{baulieu1988a,myers1990c,brandhuber1994a,piguet1995a,sadovski2017c,sadovski2018a,sadovski2018b}. A particularly convenient one is the (anti-)self-dual Landau ((A)SDL) conditions % chktex 36
\begin{subequations}\label{eq:asdlg}
  \begin{align}
    d \star A    & = 0 \;, \\
    d \star \psi & = 0 \;, \\
    F^\pm        & = 0 \,,
  \end{align}
\end{subequations}
where $ F^\pm \equiv F \pm \star F $. This choice localizes the path integral on the very well-behaved $SU(N)$ YM (anti-)instanton moduli, where the Donaldson invariants were originally formulated via intersection theory~\cite{donaldson1983a,donaldson1990a,donaldson1990b}. The (A)SDL gauge results in a very strong set of Ward identities: QTYM theories are renormalizable to all orders in perturbation theory with only one independent, and non-physical, renormalization parameter~\cite{sadovski2017c}, and; all their connected $n$-point Green functions are tree-level exact~\cite{sadovski2018a}. Clearly, this gauge makes evident that QTYM are trivial as local QFTs. And, due to their lack of loop corrections, their quantum observables are the classical observables in~\eqref{eq:tym-observables}: the Donaldson polynomials. In this sense, QTYM and classical TYM represent the same physical theories. % chktex 36

A final unusual feature worth notice relates to the so-called \textit{Gribov ambiguities}~\cite{gribov1978a}. These ambiguities are present in every non-Abelian gauge theories, including TYM, due to the non-triviality of the gauge moduli bundle $\mathcal{G} \hookrightarrow \mathcal{A} \rightarrow \mathcal{M}$~\cite{singer1978a}. However, different from most non-Abelian gauge theories, TYM theories are tree-level exact. Therefore, a refinement of the gauge fixing procedure \textit{\`a la} Gribov-Zwanzinger, for instance, cannot yield any interesting new physics. In this sense, we say, their Gribov problems are trivial~\cite{sadovski2020a}.

\end{document}
